This document contains all definitions required to score centralised
cross-country competitions for both hang-gliding and paragliding. Its main
purpose is to serve as an addendum to sections 7A and 7B of the FAI sporting
code. Additionally, it should serve as an educational tool for all parties
involved in such competitions, as a reference for the implementation of scoring
systems, as well as a basis for future improvements and modifications.

\subsection{Scope}
The document’s scope is restricted to scoring of FAI Category 1 cross-country
competitions for hang-gliding and paragliding: World and Continental
championships in both sports. CIVL’s rule setting targets these competitions
exclusively, whereas organisers of FAI Category 2 competitions as well as
non-sanctioned competitions are free to score their competitions however they
like. Most of them do follow CIVL’s lead, though, so this document should also
cover the majority of Category 2 events as well.

\subsection{Sources}
The remainder of this document is based on:
\begin{itemize}
\item “The GAP Guide” (2011 edition)
\item FAI Sporting Code Section 7A for Hang-Gliding (2013 edition)
\item FAI Sporting Code Section 7B for Paragliding (2013 edition)
\item The scoring implementation within CIVL’s scoring software, FS (aka “FScomp”)
\item Appendix C of the Paragliding World Cup Association’s 2013 Competition Rules
\end{itemize}

\subsection{Changes from previous edition}
Changes from the previous edition of this document are underlined (a few words)
or marked with a vertical sidebar (whole paragraph).

\subsection{Differences between Hang-Gliding and Paragliding}
Initially, both hang-gliding and paragliding competitions used the same system
for scoring, generally known as “GAP”. But over time, through their two
separate sporting codes, the two disciplines introduced more and more changes
that would only apply to one, but not the other. This mainly in non-standard
situations such as stopped tasks, pilots landing just short of goal, or pilots
crossing the start line too early. Where not explicitly stated otherwise, the
contents of this document always apply to both disciplines. Definitions
applying only to one, but not the other, are clearly marked as such.

\begin{hg}
Text marked in blue applies exclusively to hang-gliding.
\end{hg}

\begin{pg}
Text marked in orange applies exclusively to paragliding.
\end{pg}
